Categorical and gradient aspects of wordlikeness\thanks{Thanks to Gene Buckley, Charles Yang, and audiences at NELS 43 for helpful comments.}

### 2013-01-28 4.tex
Revisions! Lots of 'em.

It is a curious fact that many constraints on underlying representations are elegantly described using the units of prosody.
In some cases, these are part of the system of contrast, such as long segments in moraic theory \citep[e.g.,][]{Davis1999}. For instance, Latin has constrastive consonant and vowel length (e.g., \emph{anus} `ring' vs.~\emph{annus} `year', \emph{os} `bone' vs.~\emph{ōs} `mouth'), but there is no vowel length contrast before geminate consonants. 
However, constraints on underlying representations may also involve references to non-contrastive prosodic structures, such as syllables (e.g., \citealt{Hooper1973}, \citealt{Kahn1976}).\footnote{
    See \citealt{Blevins1995} for the claim that syllable structure is universally non-contrastive, and \citealt{Elfner2006} for arguments that a putative counterexample is the result of an underlying moraic, not syllabic, contrast.}
For instance, as \citeauthor{Haugen1956} first notes, the constraints on word-medial consonant clusters are not arbitrary in prosodic terms:

\begin{example}[\textsc{Medial Cluster Law} (after \citealt{Haugen1956})]
\label{mcl}
A medial cluster consists maximally of a well-formed medial coda and a well-formed medial onset
\end{example}

\noindent
It is not necessary for predictable prosodic structure to be present in underlying form to achieve this effect, contrary to what has sometimes been stated \citep[e.g.,][255]{A74}.
All that is necesary is that no phonological processes intervene between underlying representation and the surface form of medial clusters.
This is little more than the definition of a ``faithful'' cluster.
Under the further assumption that underlying and surface representations are as similar as phonological processes allow (e.g., \citealt[205f.]{Dell1973}, \citealt[28f.]{Stampe1973}), faithful realization is the null hypothesis.

In a study of English word-medial consonant clusters, \citet{Pierrehumbert1994} notes that the vast majority of ``possible clusters'', those which can be parsed into coda and onset in accordance with the \textsc{Medial Cluster Law}, are unattested.
\citeauthor{Pierrehumbert1994} argues that this can be accounted for by positing static co-occurrence restrictions---unrelated to any phonological alternation in English.
This study argues that well-known phonological processes, but not the static constraints proposed by \citeauthor{Pierrehumbert1994}, impose restrictions on this inventory.
%\citeauthor{Pierrehumbert1994}'s static constraints are not statistically valid

@@

English syllable contact clusters have been used, by \citeauthor{Pierrehumbert1994} and others, to argue for the necessity of admitting static phonotactic constraints, and to illustrate proposals for the architecture of the phonotactic system. 
However, there are a number of reasons to reconsider this domain.

@@

\citeauthor{Pierrehumbert1994} does not fully consider the effects of phonological processes targeting word-medial clusters in English, and consequently some of the static constraints she proposes may be related to English morphophonemics. 
For instance, she writes that ``nasal-stop sequences agree in labiality'' (175), and attributes this the status of a static constraint. 
However, this generalization is just a narrower form of the one derived by \textsc{Nasal Place Assimilation} (see \S\ref{npa}).
Other derived constraints go unnoticed: for instance, she does not discuss the highly reliable tendency of obstruent-obstruent clusters to agree in voicing (see \S\ref{ova}). 
The evaluation below considers the role of both types of constraints on the English lexicon on an equal footing.

@@

\citet{Pierrehumbert1994} infers static constraints from near-exceptionless gaps in the lexicon, but little effort is made to show that the patterns of lexical underrepresentation are not due to chance. 
Consquently, it is possible to suggest that some of these gaps 
are accidental rather than structural in nature.
%\citep{Fischer-Jorgensen1952,Vogt1954}. 
This is made all the more likely given the tendency of segment and cluster frequency distributions to be highly skewed \citep[e.g.,][]{Pande2010,Sigurd1968,Tambovtsev2007,Weiss1961}. 
Furthermore, \citeauthor{Pierrehumbert1994} considers only triconsonantal clusters, but medial clusters may be as short as two consonants, as in \emph{a}[n.t]\emph{ics}, or as long as four, as in \emph{mi}[n.str]\emph{el}, and \citeauthor{Pierrehumbert1994} offers no justification for a narrower focus.
If this has any effect on the distribution of clusters, it would be to impose further sparsity.

@@

Many components of \citeauthor{Pierrehumbert1994}'s study cannot be directly replicated. \citeauthor{Pierrehumbert1994} focuses her attention on those words she judges to be ``morphologically simple'' and ``reasonably familiar'', but this list has neither been published nor circulated.
It is not clear that the judgements (as well as cognitive limitations) of concerned parties should be granted evidential status in the first place, given the potential for implicit bias; 
\citet{Labov1975} for instance
argues for what he calls the \emph{Experimenter Principle}, 
that the intuitions of all those familiar with the formal matters should be excluded.
Otherwise-unmotivated morphological junctures are sometimes posited simply to preserve phonological or phonotactic generalizations; 
%The same practice is sometimes applied to preserve phonological generalizations; for instance, otherwise unmotivated morphological boundaries are used in \emph{SPE} to simplify principles of stress assignment. 
this is done by \citet{SPE}, for instance, to simplify principles of English stress assignment, and \citet[546]{Rice2009d} analyses many words in Slave as compounds simply because they contain consonant clusters that rarely occur morph-internally. 
%There is some evidence that infants \citep{Mattys2001b} and adults \citep{Brown1956,Hay2004a,McQueen1998b,Norris1997} use this heuristic to segment fluent speech in experimental settings. 
Applied indiscriminately, however, such a heuristic trivializes both morphological segmentation and phonotactic generalization.
For these reasons, the wordlist used in this study is derived from a publicly available database, and no experimenter intuitions are used for establishing morphological boundaries.

%Such reasoning is found in \citeauthor{Pierrehumbert1994}'s discussion of the word \emph{a}[nt.l]\emph{er}, however: this contains a cluster which violates her constraint on coda coronal obstruents, and therefore must be a complex word, she argues (see \citealt[164]{Borowsky1989} for additional exceptions to this generalization).

@@

After constructing a sample of syllable contact clusters in English simplex words, this data is used to evaluate the coverage of static and phonologically derived constraints.

@@

\subsubsection{Materials and procedure}

Following \citet[ chap.~8]{Duanmu2009} and \citet[ chap.~3]{Hammond1999a}, 
a wordlist is generated using the English portion of the CELEX database \citep{CELEX}. 
Only words marked in CELEX as ``monomorphemic'' are used, and all words 
labeled non-native are excluded.
This stringent criterion for inclusion eliminates many of the exceptions noted by \citeauthor{Duanmu2009} or \citeauthor{Hammond1999a} in their studies. For instance, nearly all the exceptions to \textsc{Obstruent Voice Assimilation} (see \S\ref{ova} below) noted by \citet[74]{Hammond1999a} are 
either complex words (e.g., \emph{jurisdiction}, \emph{madcap}, \emph{tadpole}, \emph{scapegoat}, \emph{magpie}) or non-native (e.g., \emph{vodka}, \emph{smorgasbord}).

Unlike prior studies, words which appear to consist of a Latinate prefix and a bound stem (e.g., \emph{inspect}, \emph{excrete}) are also excluded from the sample, as there is broad formal and experimental support for the claim that these words are complex.
This assumption allows for numerous simplifications to accounts of the phonological shape of these forms. 
For instance, \citet[11f.]{Aronoff1976} observes that Latinate forms which share the same bound stem also share irregular allomorphs of that stem under derivation.

@@

\noindent
\citeauthor{Aronoff1976} takes this to be evidence that \emph{adhere} and \emph{cohere}, for instance, share a bound stem.
There is also an interaction between Latinate prefixes on verbs and the makeup of the verbal complement. 
Latinate verbs do not generally allow ditransitive, verb participle, or adjectival resultative constructions, all of which are acceptable with similar Anglo-Saxon verbs \citep{Gropen1989,Harley2009}.
%\citep[see also][]{Gropen1989,Coppock2008}.

\begin{example}[Syntactic restrictions on Latinate verbs]
\label{harley}
\begin{tabular}{l l l l@{} l}
a. & {show him the painting} & \alt{} & * & {exhibit him the painting} \\
b. & {drink himself stupid}  & \alt{} & * & {imbibe himself stupid}    \\
%c. & {show it off}           & \alt{} & * & {exhibit it off}           \\
c. & {break it off}           & \alt{} & * & {terminate it off}        \\
\end{tabular}
\end{example}

Lexical decision experiments also suggest that Latinate prefixed forms are complex. \citet{Taft1975,Taft1976} and \citet{Taft1986} find that nonce words like \emph{*re-sert}, which appear to be composed of a prefix and a bound stem, take longer to reject that non-words which lack apparent morphological structure, such as *\emph{refant}. 
Bound stems also show frequency effects which are independent of whole word frequency \citep{Taft1979,Taft2006}. 
Finally, \citet{Emmorey1989} and \citet{Forster2000} report facilitative priming between pairs like \emph{permit}-\emph{submit}, an effect thought to indicate morphological relatedness.

@@

The filters discussed above result in a list of 6,619 simplex words.
The full set of clusters and their frequencies are given in Appendix \ref{clusters}.
The CELEX transcriptions of these words were then syllabified and phonologized using a technique described in Appendix \ref{syllabification}. 
This produces 23 unique medial codas and 40 unique medial onsets. 
Of the 920 ($= 21 \times 40$) medial clusters that would result from free combination, 174 (19\%) are attested.

@@

So as to account for the 81\% of ``possible'' but unattested clusters, \citet{Pierrehumbert1994} proposes three constraints on English medial clusters. These constraints are static, in that they have no analogues in the morphophonemics of English.

\paragraph{Dorsal-labial clusters} 
\citet[173]{Pierrehumbert1994} writes that ``velar obstruents occurred only before coronals in the clusters studied, never before labials or other velars'', noting that velar-velar clusters are independently excluded by constraints on geminates (see \S\ref{degemx}). 
Two-consonant velar-labial clusters are found in words like \emph{a}[k.m]\emph{e}, \emph{ru}[ɡ.b]\emph{y}, or \emph{pi}[ɡ.m]\emph{ent} occur. 
Table \ref{dltab} shows that velar-labial clusters are somewhat less likely to occur than  are velar-coronal clusters (e.g., \emph{ve}[k.t]\emph{or}), but such underrepresentation is not unlikely to occur by chance ($p = .106$).

@@

\paragraph{Coronal obstruent codas} 
\citet[175]{Pierrehumbert1994} claims that ``clusters with a coronal obstruent in the coda do not occur'', while noting exceptions like \emph{a}[nt.l]\emph{er}, \emph{ke}[s.tr]\emph{el} and \emph{oi}[nt.m]\emph{ent}. 
Table \ref{cctab} reveals that coda coronal obstruent clusters are not significantly less likely to occur than non-coronal obstruent clusters (e.g., \emph{re}[p.t]\emph{ile}). 
The same is true if attention is restricted to triconsonantal clusters ($p = .129$).

@@

\paragraph{ABA clusters} 
\citet[176]{Pierrehumbert1994} notes the ``lack of clusters with identical first and third elements'', ignoring presence or absence of voicing. 
Despite the fact that there are no exceptions to this generalization, these \textsc{ABA} clusters are not significantly less common than other triconsonantal and quadraconsonantal clusters (Table \ref{abatab}).

@@

\paragraph{Summary} 
There is no statistical support for \citeauthor{Pierrehumbert1994}'s static constraints.

@@

In \emph{The Sound Pattern of English} (henceforth, \emph{SPE}), \citet{SPE} describe three phonological processes which target medial consonant clusters. As will be shown, the neutralizing effect of these three processes has a major influence on the cluster inventory.

@@

Voice assimilation alternations are evidenced by the non-syllabic allomorphs of the regular past (e.g., \emph{nap}[t] $\sim$ \emph{nab}[d]) and noun plural (e.g., \emph{lap}[s] $\sim$ \emph{lab}[z]), which take the voicing specification of a preceding obstruent.\footnote{Underlying /-d, -z/ are assumed here (e.g., \citealt{Anderson1973a}, \citealt[284f.]{Bakovic2005b}, \citealt{Basboll1972}, \citealt[210]{SPE}, \citealt[282]{Hockett1958}, \citealt[102]{Pinker1988}, \citealt{Shibatani1972}); alternative analyses are put forth by \citet[210f.]{LANGUAGE}, \citet[135]{Borowsky1986}, \citet{Hoard1971}, \citet{Kiparsky1985}, \citet{Lightner1970}, \citet{Luelsdorff1969}, \citet{Miner1975}, \citet[426]{Nida1948}, and \citet{Zwicky1975}.}
%Bizarely, \citet[208]{Wetzels2001}, cite English as an example of a language without ``general devoicing or assimilatory effects''.

@@

\noindent
\citet{Pierrehumbert1994} does not discuss a constraint against adjacent obstruents disagreeing in voice. 
As shown in Table \ref{ovatab} however, the vast majority of medial obstruents clusters in simplex words are either uniformly voiced, as in \emph{hu}[z.b]\emph{and}, or uniformly voiceless, as in or \emph{rha}[p.s]\emph{osdy}. Hetero-voiced clusters, like those in \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}, are rarer than would be expected from chance.

@@

\textsc{Nasal Place Assimilation} (e.g., \citealt[65f.]{Borowsky1986}, \emph{SPE}:85, \citealt[62]{Halle1985a}) 
%\citealt[269]{Fudge1969}) 
permits [ŋ] to be described as a pure allophone of /n/, and accounts for certain patterns in Latinate prefix allomorphy.

@@

\noindent
As shown in Table \ref{npatab}, virtually all clusters consisting of a nasal coda followed by a homorganic obstruent (e.g., \emph{pi}[m.p]\emph{le}, \emph{sta}[n.z]\emph{a}, \emph{mo}[ŋ.k]\emph{ey}) are attested. As \citet[175]{Pierrehumbert1994} observes, heterorganic clusters, like those \emph{pli}[m.s]\emph{oll} or \emph{scri}[m.ʃ]\emph{aw}, do occur, but they are significantly more rare.

@@

\paragraph{Degemination} 
The final alternation found in English medial clusters is the simplification of geminates which is characteristic of ``level I'' morphology, and is found in the irregular /-t/ past tense (e.g., \emph{bend}/\emph{ben}[t], \emph{build}/\emph{buil}[t]), in \emph{-ly} deadjectival derivatives (e.g., \emph{norma}[l]\emph{y}, cf. \emph{calm}[l]\emph{y}), and Latinate prefix allomorphy (\emph{SPE}:148; \citealt[102]{Borowsky1986}). \textsc{Degemination} is formalized here as a rule deleting the first of two segments agreeing on all feature values (except voice, possibly).
\label{degemx}

@@

\noindent
As can be seen from Table \ref{degemtab}, the sample includes no sequences of identical segments, or of identical segments differing only in voice, something highly unlikely to arise by chance.
Stampean occultation by \textsc{Degemination} provides a natural explanation for this gap.
It is informative to compare this to the constraint against ``ABA'' clusters proposed by \citeauthor{Pierrehumbert1994}: both are exceptionless, but only the former is unlikely to arise by chance.

@@

Current computational models of phonotactic knowledge ``rate'' possible clusters by producing a numerical value for each cluster. 
To apply these models to predicting the cluster inventory, these numerical values must be transformed into a categorical prediction of either attestation or non-attestation.
This is accomplished with a soft-margin support vector machine \citep{Cortes1995} with a linear kernel, which optimally splits the ratings distribution into a set of attested and unattested clusters.
This classifier does not correspond to any component of a cognitively plausible model of phonotactic learning: it simply represents an upper bound for predicting the cluster inventory from positive data.

Four metrics are used to score the models. 
\emph{Accuracy} represents the probability that a cluster is correctly classified as attested or unattested.
Two additional metrics break down accuracy into its constituent parts;
\emph{precision} represents the probability that a cluster which is predicted to be unattested is in fact unattested, and \emph{recall} is the probability that an unattested cluster is predicted as such.
It is possible to increase precision at the expense of recall, by predicting non-attestation for a greater number of clusters, or to maximize recall at the expense of precision by predicting all clusters to be unattested. 
$F_1$, the harmonic mean of these two measures, is a standard metric for quantifying this tradeoff; increases in either precision or recall produce a  non-linear increase in $F_1$. The results are summarized in Table \ref{cmresults}.

\begin{table}
\centering
\begin{tabular}{l | rrrr}
\toprule
                    & accuracy & precision & recall & $F_1$ \\
\midrule
Baseline            & 0.812    & 0.812     & 1.000  & 0.896 \\
Expected frequency  & 0.836    & 0.845     & 0.984  & 0.909 \\
Derived constraints & 0.838    & 0.840     & 0.997  & 0.912 \\
DC \& EF            & 0.861    & 0.868     & 0.988  & 0.924 \\
\citet{Hayes2008a}  & 0.835    & 0.942     & 0.854  & 0.896 \\
\bottomrule
\end{tabular}
\caption{Results for the cluster classification task}
\label{cmresults}
\end{table}
%; a combination of derived constraints and expected frequency (``DC \& EF'') produces the highest accuracy and $F_1$.}

>>>>

It is indisputable that many phonotactic restrictions are easily described with reference to prosodic primitives.
In some cases, prosodic factors constrain the system of contrast.
For instance, Latin has contrastive vowel and consonant length (e.g., \emph{os} `bone' vs.~\emph{ōs} `mouth', \emph{anus} `ring' vs.~\emph{annus} `year'), but the latter contrast is suspended in codas preceded by a diphthong or long monophthong; a syllable may contain a long vowel, or be checked by the first half a geminate consonant, but not both.
However, constraints on underlying representations may also involve references to non-contrastive prosodic structures such as the syllable (e.g., \citealt{Hooper1973}, \citealt{Kahn1976}).\footnote{
    See \citealt{Blevins1995} for the claim that syllable structure is universally non-contrastive, and \citealt{Elfner2006} for arguments that a putative counterexample derives from an underlying vowel length contrast.}
For instance, as noted by \citet{Haugen1956}, 
numerous restrictions on word-medial consonant clusters have a unified statement in prosodic terms:

\begin{example}[\textsc{Medial Cluster Law}]
\label{mcl}
A medial cluster can consist maximally of a well-formed medial coda and a well-formed medial onset
\end{example}

As an illustration of this tautology, consider languages like Yokuts, which forbid complex codas and complex onsets, and which enforce these restrictions in complex words via processes such as vowel epenthesis.
\citet[26f.]{Newman1944} notes that this imposes an an upper bound on the size of medial clusters: no cluster of more than two consonants can be parsed into a simple coda and simple onset.
In the case of Yokuts, it is certainly possible to state this restriction without reference to syllable structure, as *CCC, a constraint on trisyllabic clusters (e.g., \citealt[92f.]{Ettlinger2008}, \citealt[820f.]{Zuraw2003a}).
However, the aforementioned constraints on complex onsets and codas find independent motivation from the total absence of initial and final clusters in Yokuts;\footnote{
    The tendency of consonants to pattern with word boundaries and morph junctures has been long been noted \citep[e.g.,][]{Hill1954,Lass1971,Moulton1947}; \citet[24f.]{Kahn1976} takes this as evidence for the syllable.}
with these two restrictions in place, a further constraint on medial triconsonantal clusters is otiose.\footnote{
    \citet[31f.]{Cote2000} alludes to another critique of constraints like *CCC, namely that this constraint requires ``counting'': see \citealt[64f.]{Isac2008} for further discussion of the comparative merits of ``counting'' and ``grouping'' analyses in phonology.}
While the medial cluster law is certainly consistent with the hypothesis that syllable structure may be present in underlying representations (e.g., \citealt{A74}:255, \citealt{Vaux2003}), this need not be the case under Stampean occultation.
If an underlying medial consonant sequence appears on the surface, it satisfies the medial cluster law by definition.
If, however, an underlying cluster is modified by consonant deletion, coalescence, or vowel epenthesis, then it need not consist of a licit medial onset and medial coda.

\citet{Pierrehumbert1994} begins a study of English word-medial consonant clusters with a restatement of the medial cluster law:

\begin{quote}
That is, in the absence of additional provisos, any concatenation of a well-formed coda and a well-formed onset is predicted to be possible medially in a word. \citep[168]{Pierrehumbert1994}
\end{quote}

\noindent
However, \citeauthor{Pierrehumbert1994} reports that the vast majority of the ``possible'' clusters (i.e., those which conform to the medial cluster law) are in fact unattested. 
So as to acount for this, \citeauthor{Pierrehumbert1994} presents ``provisos'' in the forms of static co-occurrence restrictions, unrelated to any phonological alternation in English.\footnote{
    Given the arguments presented in the previous chapter, that not all lexical tendencies have a synchronic basis, one may question the intuition that any gaps in the English cluster inventory (or those that go beyond the medial cluster law) must be accounted for by the synchronic grammar.
    In addition to some informal statistical evidence (of the sort problematized in the previous chapter), \citeauthor{Pierrehumbert1994} administers a wordlikeness task to validate the static constraints she proposes.
    However, this experiment is of a quite informal nature and the results are given only a superficial analysis, so it is less than probative.} 
This chapter will argue, however, the static constraints proposed by \citeauthor{Pierrehumbert1994} are unnecessary, and that the only restrictions on the inventory of medial clusters in English (beyond the medial cluster law) are those which derive from well-known phonological processes.

@@

The aforementioned study by \citeauthor{Pierrehumbert1994}, as well as further investigations of this domain by \citet[ chap.~8]{Duanmu2009} and \citet[ chap.~3]{Hammond1999a}, argue for the necessity of admitting static phonotactic constraints, and illustrate proposals for the architecture of the phonotactic system.
However, there are a number of reasons to reconsider the findings of these authors in light of the proposals made in previous chapters.

@@

\citeauthor{Duanmu2009}, \citeauthor{Hammond1999a}, and \citeauthor{Pierrehumbert1994} do not generally take into account the effects of phonological processes which target medial clusters in English.
As a consequence, some of the static constraints these authors identify may be in fact the product of English morphophonemics and Stampean occultation. 
For instance, \citeauthor{Pierrehumbert1994} writes that ``nasal-stop sequences agree in labiality'' (175) and posits a static constraint to account for this fact.
However, this generalization is merely a narrower form of a restriction deriving from a process of \textsc{Nasal Place Assimilation} (see \S\ref{npa}).
Other derived constraints are simply not mentioned; for instance, \citeauthor{Pierrehumbert1994} does not discuss the highly reliable tendency of obstruent-obstruent clusters to agree in voicing (see \S\ref{ova}); while \citet{Hammond1999a} does allude to this restriction, it is dismissed in light of a few apparent counterexamples (though see \S\ref{sss:map} below).
In contrast, this chapter attempts to evaluate derived and static constraints on an equal footing.

@@

\citet{Pierrehumbert1994} infers static constraints from near-exceptionless gaps in the lexicon, but little effort is made to show that the patterns of lexical underrepresentation are not due to chance. 
Consequently, it is possible to suggest that some of these gaps are accidental rather than structural in nature.
This is made all the more likely given the tendency of segment and cluster frequency distributions to be highly skewed \citep[e.g.,][]{Pande2010,Sigurd1968,Tambovtsev2007,Weiss1961} so that it is difficult to distinguish between structural and accidental gaps.
Furthermore, \citeauthor{Pierrehumbert1994} considers only triconsonantal clusters, but medial clusters may be as short as two consonants, as in \emph{a}[n.t]\emph{ics}, or as long as four, as in \emph{mi}[n.str]\emph{el}, and no justification is given for ignoring clusters of other lengths.
If there is any effect of this focus, it is presumably to produce further sparsity in the distribution observed.

@@

Many components of \citeauthor{Pierrehumbert1994}'s study cannot be replicated.
\citeauthor{Pierrehumbert1994} limits her study to words she judges to be ``morphologically simple'' and ``reasonably familiar''; the author's sensations thereof are not replicable, nor are they available to other researchers in any form.
It has been suggested \citep[e.g.,][]{Labov1975,Schutze1996} that the sensations (as well as cognitive limitations) of concerned parties should not be granted evidential status in the first place, given the potential for implicit bias; \citeauthor{Labov1975} calls the \emph{Experimenter Principle}.

It is not uncommon for analysts to propose otherwise-unmotivated morphological junctures simply to preserve phonological or phonotactic generalizations.
This is done by \citet{SPE}, for instance, to simplify principles of English stress assignment.
Similarly, \citet[546]{Rice2009d} analyses many words in Slave as compounds simply because they contain consonant clusters that rarely occur in morph-internal contexts.
%There is some evidence that infants \citep{Mattys2001b} and adults \citep{Brown1956,Hay2004a,McQueen1998b,Norris1997} use this heuristic to segment fluent speech in experimental settings.
Applied indiscriminately, however, this heuristic trivializes both morphological segmentation and phonotactic generalization.
For these reasons, the wordlist used in this study is derived from a publicly available database, and no experimenter intuitions are used.

@@

After constructing a sample of syllable contact clusters in English simplex words, this sample is used to evaluate the coverage of static and derived constraints.

@@

\subsubsection{Materials and procedure}
\label{sss:map}

Following \citet[ chap.~8]{Duanmu2009} and \citet[ chap.~3]{Hammond1999a}, who also consider restrictions on English medial clusters, a wordlist is generated using the English portion of the CELEX database \citep{CELEX}.
Only words marked in CELEX as ``monomorphemic'' are used, and all words labeled in CELEX as non-native are excluded.\footnote{
    The first criterion results in the exclusion of proper names, which have long been noted to push the bounds of native language phonotactics \citep[e.g.,][254]{Trubetzkoy1939}.}
These more-stringent criteria exclude many words labeled exceptions in the studies by \citeauthor{Duanmu2009} or \citeauthor{Hammond1999a} in their studies.
For instance, nearly all the exceptions to \textsc{Obstruent Voice Assimilation} (see \S\ref{ova} below) noted by \citet[74]{Hammond1999a} are excluded either as complex words (e.g., \emph{jurisdiction}, \emph{madcap}, \emph{tadpole}, \emph{scapegoat}, \emph{magpie}) or non-native (e.g., \emph{vodka}, \emph{smorgasbord}).

In contrast to prior studies, these criteria also exclude words which consist of a Latinate prefix and a bound stem (e.g., \emph{inspect}, \emph{excrete}). 
While \citeauthor{Pierrehumbert1994} rejects this analysis as as unmotcontroversial, it in fact has extensive formal and experimental support.
First, Latinate prefixes simplify the statement of many morphophonemic details in English.
For instance, \citet[11f.]{Aronoff1976} observes that Latinate forms which share the same bound stem also share irregular allomorphs of that stem under derivation.

@@

\noindent
\citeauthor{Aronoff1976} takes this to be evidence that \emph{adhere} and \emph{cohere}, for instance, share a bound stem.
There is also an interaction between Latinate prefixes on verbs and the complements they select.
Latinate verbs do not generally allow ditransitive, verb participle, or adjectival resultative constructions, all of which are acceptable with similar Anglo-Saxon verbs \citep[e.g.,][]{Gropen1989,Harley2009}.

\begin{example}[Restrictions on Latinate verbal complements]
\label{harley}
\begin{tabular}{l l l l@{} l}
a. & {show him the painting} & \alt{} & * & {exhibit him the painting} \\
b. & {drink himself stupid}  & \alt{} & * & {imbibe himself stupid}    \\
c. & {break it off}          & \alt{} & * & {terminate it off}         \\
\end{tabular}
\end{example}

Lexical decision also provide evidence for the segmentation of Latinate prefixed forms.
\citet{Taft1975,Taft1976} and \citet{Taft1986} find that nonce words like \emph{*re-sert}, which appear to be composed of a prefix and a bound stem, take longer to reject that non-words which lack apparent morphological structure, such as *\emph{refant}. 
Bound stems also show frequency effects independent of whole word frequency \citep{Taft1979,Taft2006}.
Finally, \citet{Emmorey1989} and \citet{Forster2000} report facilitative priming, thought to implicating morphological relatedness, between pairs like \emph{permit}-\emph{submit}, which appear to share a bound stem.

@@

Filtering the CELEX data according to the above criteria results in a list of 6,619 simplex words.
The full set of clusters and their frequencies are listed in Appendix \ref{clusters}.
The CELEX transcriptions of these words are then syllabified and phonologized using the same procedure as used in Chapter \ref{gradience} and described in Appendix \ref{syllabification}. 
In all, the sample contains 23 different medial coda and 40 different medial onsets.
Of the 920 ($= 21 \times 40$) medial clusters that would result from free combination of medial coda and medial onset, 174 (19\%) are attested.

@@

To account for the 81\% of ``possible'' but unattested clusters, \citet{Pierrehumbert1994} proposes three static constraints on English medial clusters.

\paragraph{Dorsal-labial clusters}
\citet[173]{Pierrehumbert1994} writes that ``velar obstruents occurred only before coronals in the clusters studied, never before labials or other velars'', while noting that absence of velar-velar clusters is expected due to a separate constraint on geminate clusters (see \S\ref{degemx}).
However, biliteral velar-labial clusters are found in words such as \emph{a}[k.m]\emph{e}, \emph{ru}[ɡ.b]\emph{y}, or \emph{pi}[ɡ.m]\emph{ent}.
Velar-labial clusters are somewhat less common than velar-coronal clusters (e.g., \emph{ve}[k.t]\emph{or}), but such underrepresentation is not unlikely to occur by chance according to the Fisher exact test (Table \ref{dltab}).

@@

\paragraph{Coronal obstruent codas}
\citet[175]{Pierrehumbert1994} claims that ``clusters with a coronal obstruent in the coda do not occur'', but at the same time observes exceptions like \emph{a}[nt.l]\emph{er}, \emph{ke}[s.tr]\emph{el} and \emph{oi}[nt.m]\emph{ent}.
In the CELEX sample (Table \ref{cctab}), coda coronal obstruent clusters are not significantly less likely to occur than non-coronal obstruent clusters (e.g., \emph{re}[p.t]\emph{ile}).
While not shown in tabular form, the same is true if attention is restricted to triconsonantal clusters ($p = .129$).

@@

\paragraph{ABA clusters} 
\citet[176]{Pierrehumbert1994} observes a ``lack of clusters with identical first and third elements'', ignoring presence or absence of voicing. 
Despite the fact that there are no exceptions to this generalization, these \textsc{ABA} clusters are not significantly less common than any other triconsonantal and quadraconsonantal clusters (Table \ref{abatab}).

@@

\paragraph{Summary} 
There is no statistical support for any of \citeauthor{Pierrehumbert1994}'s static constraints.

@@

In \emph{SPE}, \citet{SPE} describe three phonological processes which target medial consonant clusters. 
As will be shown, these three processes have a profound influence on the English cluster inventory.

@@

Voice assimilation alternations are evidenced by the non-syllabic allomorphs of the regular past (e.g., \emph{nap}[t]-\emph{nab}[d]) and noun plural (e.g., \emph{lap}[s]-\emph{lab}[z]), which take the voicing specification of a preceding obstruent;\footnote{
    Underlying /-d, -z/ are assumed here (e.g., \citealt{Anderson1973a}, \citealt[284f.]{Bakovic2005b}, \citealt{Basboll1972}, \citealt[210]{SPE}, \citealt[282]{Hockett1958}, \citealt[102]{Pinker1988}, \citealt{Shibatani1972}); alternative analyses are put forth by \citet[210f.]{LANGUAGE}, \citet[135]{Borowsky1986}, \citet{Hoard1971}, \citet{Kiparsky1985}, \citet{Lightner1970}, \citet{Luelsdorff1969}, \citet{Miner1975}, \citet[426]{Nida1948}, and \citet{Zwicky1975}.} 
voice assimilation is also claimed to operate across prefix and compound junctures \citep{Davidsen-Nielsen1974}.
%Bizarely, \citet[208]{Wetzels2001}, cite English as an example of a language without ``general devoicing or assimilatory effects''.

@@

\noindent
\citet{Pierrehumbert1994} does not discuss a constraint against adjacent obstruents disagreeing in voice. 
However, the vast majority of medial obstruents clusters in simplex words are either uniformly voiced, as in \emph{hu}[z.b]\emph{and}, or uniformly voiceless, as in or \emph{rha}[p.s]\emph{osdy} (Table \ref{ovatab}).
Hetero-voiced clusters, like those in \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}, are far rarer than would be expected from chance.

@@

\textsc{Nasal Place Assimilation} (e.g., \citealt[65f.]{Borowsky1986}, \emph{SPE}:85, \citealt[62]{Halle1985a})
%\citealt[269]{Fudge1969})
permits [ŋ] to be described as an allophone of /n/ (see \S\ref{s:phonologization}), and furthermore accounts for allomorphy in certain Latinate prefixes.

@@

\noindent
Virtually all clusters consisting of a nasal coda followed by a homorganic obstruent (e.g., \emph{pi}[m.p]\emph{le}, \emph{sta}[n.z]\emph{a}, \emph{mo}[ŋ.k]\emph{ey}) are attested (Table \ref{npatab}).
As \citet[175]{Pierrehumbert1994} observes, heterorganic clusters, like those \emph{pli}[m.s]\emph{oll} or \emph{scri}[m.ʃ]\emph{aw}, do occur, but in this sample they are significantly more rare.

@@

\paragraph{Degemination} 
The final alternation found in English medial clusters is the simplification of geminates which is characteristic of ``level I'' morphology, and is found in the irregular /-t/ past tense (e.g., \emph{bend}/\emph{ben}[t], \emph{build}/\emph{buil}[t]), in \emph{-ly} deadjectival derivatives (e.g., \emph{norma}[l]\emph{y}, cf.~\emph{calm}[l]\emph{y}), and Latinate prefix allomorphy (\citealt[102]{Borowsky1986}, \emph{SPE}:148).
\textsc{Degemination} is formalized here as a rule deleting the first of two segments agreeing on all feature values (except for voice, possibly).
\label{degemx}

@@

\noindent
The sample contains no sequences of identical segments, or of identical segments differing only in voice, something highly unlikely to arise by chance (Table \ref{degemtab}); \textsc{Degemination} and Stampean occultation provide a natural explanation for this gap.
It is interesting to compare the absense of geminates to the constraint against ``ABA'' clusters proposed by \citeauthor{Pierrehumbert1994} in this regardl: both are exceptionless, but only the former imposes a lexical tendency unlikely to arise by chance.

@@

Current computational models of phonotactic knowledge can ``rate'' possible clusters, assigning a numerical wellformedness score to any input.
These models can be applied to a task of predicting which clusters are and are not attested by transforming these numerical values into a categorical prediction of either attestation or non-attestation.
This is accomplished here with a soft-margin support vector machine \citep{Cortes1995} with a linear kernel, which attempts to find a single optimal numerical value about which to split attested and unattested clusters.
This classifier is not intended to correspond to any component of a cognitively plausible model of phonotactic learning: it simply represents an upper bound for predicting the cluster inventory from positive data.

The models are scored using a ``leave-one-out'' scheme, in which each observation (a cluster) is scored using a model trained on all other observations.
Four metrics are used to evaluate model fit.
\emph{Accuracy} represents the probability that a cluster is correctly classified as attested or unattested.
Two additional metrics break down accuracy into constituent parts;
\emph{precision} represents the probability that a cluster which is predicted to be unattested is in fact unattested, and \emph{recall} is the probability that an unattested cluster is predicted as such.
It is possible to increase precision at the expense of recall, by predicting non-attestation for a greater number of clusters, or to maximize recall at the expense of precision by predicting all clusters to be unattested.
$F_1$, the harmonic mean of these two measures, is a standard metric for quantifying this tradeoff; any increase in either precision or recall will result in an increase in $F_1$.
The results are summarized in Table \ref{cmresults}.

\begin{table}
\centering
\begin{tabular}{l | rrrr}
\toprule
                    & accuracy & precision & recall & $F_1$ \\
\midrule
Baseline            & 0.812    & 0.812     & 1.000  & 0.896 \\
Expected frequency  & 0.835    & 0.837     & 0.960  & 0.894 \\
Derived constraints & 0.838    & 0.835     & 0.967  & 0.897 \\
DC \& EF            & 0.861    & 0.866     & 0.969  & 0.914 \\
\citet{Hayes2008a}  & 0.835    & 0.964     & 0.833  & 0.894 \\
\bottomrule
\end{tabular}
\caption{Results for the cluster classification task}
\label{cmresults}
\end{table}
%; a combination of derived constraints and expected frequency (``DC \& EF'') produces the highest accuracy and $F_1$.}
